\documentclass [11pt, proquest] {uwthesis}[03/03/15] %[2021/06/15]
\usepackage{graphicx}
\usepackage[sort&compress,numbers]{natbib}
\usepackage{multirow}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[breaklinks,hidelinks]{hyperref}
\usepackage{amsmath,amssymb,amsfonts,amsthm,epsfig,epstopdf,titling,url}
\usepackage{enumerate}
\usepackage{algorithm,algorithmic}
\usepackage{{booktabs}}
% \usepackage{tikz}
% \usepackage{tabularx}
% \usepackage{booktabs} % for better table lines
% \usepackage{array} 
% \renewcommand\tabularxcolumn[1]{>{\raggedright\arraybackslash}p{#1}}
\usepackage{enumitem}


\usepackage[numbers]{natbib}
\newenvironment{packed_enum}{
\begin{enumerate}[a)]
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
}{\end{enumerate}}

\newenvironment{packed_enum1}{
\begin{enumerate}[1.]
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
}{\end{enumerate}}

\newenvironment{packed_enum_par}{
\begin{enumerate}[1)]
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
}{\end{enumerate}}

\newenvironment{packed_e}{
\begin{enumerate}
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
}{\end{enumerate}}

\newenvironment{packed_itemize}{
\begin{itemize}
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
}{\end{itemize}}

% ==========   Local defs and mods

% ----------- definitions format
\theoremstyle{definition}
\newtheorem{defn}{Definition}
\newtheorem{conj}{Conjecture}[section]
\newtheorem{exmp}{Example}[section]



% ================================
% DOCUMENT
% ================================
\begin{document}

\prelimpages
% \raggedbottom


% ---------------------------

\Title{Project TLDR: Standalone desktop application for question answering and summarization using resource-efficient LLMs}

\Author{Manu Hegde}
\Year{2025}
\Program{Computer Science and Software Engineering}

\Chair{Erika Parsons}{}{School of Science, Technology, Engineering \& Mathematics}
\Signature{Michael Stiber}
\Signature{Shane Steinert-Threlkeld}

\copyrightpage

\titlepage  


\setcounter{page}{-1}
\abstract{

This project presents the design and development of a standalone desktop application for offline question answering and summarization over a user-provided document corpus, using resource-efficient large language models (LLMs). Targeted for Apple’s M1/M2 hardware, the application leverages on-device computation via the Apple Neural Engine (ANE) and Metal shaders, exploring the use of the NPU beyond traditional CoreML applications. The application addresses key concerns around data privacy, resource efficiency, and accessibility. Unlike cloud-based services that require constant internet access and raise privacy risks, this application offers a secure, local alternative optimized for researchers and students. It features a graphical interface and supports retrieval-augmented generation (RAG) over the user's corpus, all while utilizing only a fraction of system resources to support seamless multitasking. Evaluation is conducted using both functional metrics (e.g., BERTScore against ChatGPT outputs) and non-functional metrics (e.g., memory and CPU usage). The result is a practical, efficient application that enables interaction with large academic corpora while preserving system responsiveness and data confidentiality.


}

%
% ----- contents & etc.
%
\tableofcontents
\listoffigures
\listoftables  

%
% ----- acknowledgments
%
\acknowledgments{% \vskip2pc
  % {\narrower\noindent
I would like to express my gratitude to Prof. Erika Parsons for all the valuable guidance and help during this work. Furthermore, I sincerely thank Prof. Steinert-Threlkeld and Prof. Stiber for accepting my request to be on the committee for this thesis and for providing precise feedback.
  % \par}
}
%
% end of the preliminary pages
 
 
 
%
% ==========      Text pages
%
\textpages
% ============================================================
%
%                   CHAPTER 1: INTRODUCTION
%
% =============================================================
\input{1_Introduction.tex}

% ============================================================
%
%                   CHAPTER 2: THEORETICAL BACKGROUND
%
% =============================================================
\input{2_TheoreticalBackground.tex}

% ============================================================
%
%                   CHAPTER 3: RELATED WORK
%
% =============================================================
\input{3_RelatedWork.tex}

% ============================================================
%
%                   CHAPTER 4: METHODS
%
% =============================================================
\input{4_Methodology.tex}

% ============================================================
%
%                   CHAPTER 5: RESULTS
%
% =============================================================
\input{5_Experimentation.tex}

% ============================================================
%
%                   CHAPTER 6: CONCLUSION
%
% =============================================================
\input{6_Conclusion.tex}

\printendnotes

%
% ==========   Bibliography
%
% \nocite{*}   % include everything in the uwthesis.bib file
\bibliographystyle{unsrt}
\bibliography{uwthesis}
%
% ==========   Appendices
%

\end{document}