\documentclass [11pt, proquest] {uwthesis}[03/03/15] %[2021/06/15]
\usepackage{graphicx}
\usepackage[sort&compress,numbers]{natbib}
\usepackage{multirow}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[breaklinks,hidelinks]{hyperref}
\usepackage{amsmath,amssymb,amsfonts,amsthm,epsfig,epstopdf,titling,url}
\usepackage{enumerate}
\usepackage{algorithm,algorithmic}
\usepackage{{booktabs}}

\newenvironment{packed_enum}{
\begin{enumerate}[a)]
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
}{\end{enumerate}}

\newenvironment{packed_enum1}{
\begin{enumerate}[1.]
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
}{\end{enumerate}}

\newenvironment{packed_enum_par}{
\begin{enumerate}[1)]
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
}{\end{enumerate}}

\newenvironment{packed_e}{
\begin{enumerate}
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
}{\end{enumerate}}

\newenvironment{packed_itemize}{
\begin{itemize}
  \setlength{\itemsep}{1pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
}{\end{itemize}}

% ==========   Local defs and mods

% ----------- definitions format
\theoremstyle{definition}
\newtheorem{defn}{Definition}
\newtheorem{conj}{Conjecture}[section]
\newtheorem{exmp}{Example}[section]



% ================================
% DOCUMENT
% ================================
\begin{document}

\prelimpages
% \raggedbottom


% ---------------------------

\Title{Project TLDR: Standalone desktop application for question answering and summarization using resource-efficient LLMs}

\Author{Manu Hegde}
\Year{2025}
\Program{Computer Science and Software Engineering}

\Chair{Erika Parsons}{}{School of Science, Technology, Engineering \& Mathematics}
\Signature{Michael Stiber}
\Signature{Shane Steinert-Threlkeld}

\copyrightpage

\titlepage  


\setcounter{page}{-1}
\abstract{

This project presents the design and development of a standalone desktop application that enables offline question answering and summarization over a user-provided document corpus using resource-efficient large language models (LLMs). Targeted to run on Apple's M1 / M2 hardware, it leverages on-device processing through Apple's neural engine (ANE) and Apple metal shaders (GPU). The application aims to address growing concerns about data privacy, resource consumption, and user accessibility. Unlike cloud-based tools, which require constant internet connectivity and expose sensitive data to third-party servers, this application provides a secure, localized alternative optimized for academic researchers and students. Core capabilities include a graphical interface and retrieval-augmented generation (RAG) over user specified corpus, while leveraging only a fraction of system resources to allow for seamless multi-tasking. Evaluation is based on both functional quality (e.g., BERTScore comparison to ChatGPT) and nonfunctional metrics (e.g., memory, CPU usage). The outcome is a practical and efficient tool that allows users to interact with large academic corpora while maintaining system responsiveness and data confidentiality.


}

%
% ----- contents & etc.
%
\tableofcontents
\listoffigures
\listoftables  

%
% ----- acknowledgments
%
\acknowledgments{% \vskip2pc
  % {\narrower\noindent
I would like to express my gratitude to Prof. Erika Parsons for all the valuable guidance and help during this work. Furthermore, I sincerely thank Prof. Steinert-Threlkeld and Prof. Stiber for accepting my request to be on the committee for this thesis and for providing precise feedback.
  % \par}
}
%
% end of the preliminary pages
 
 
 
%
% ==========      Text pages
%
\textpages
% ============================================================
%
%                   CHAPTER 1: INTRODUCTION
%
% =============================================================
\input{1_Introduction.tex}

% ============================================================
%
%                   CHAPTER 2: THEORETICAL BACKGROUND
%
% =============================================================
\input{2_TheoreticalBackground.tex}

% ============================================================
%
%                   CHAPTER 3: RELATED WORK
%
% =============================================================
\input{3_RelatedWork.tex}

% ============================================================
%
%                   CHAPTER 4: METHODS
%
% =============================================================
\input{4_Methods.tex}

% ============================================================
%
%                   CHAPTER 5: RESULTS
%
% =============================================================
\input{5_Results.tex}

% ============================================================
%
%                   CHAPTER 6: CONCLUSION
%
% =============================================================
\input{6_Conclusion.tex}

\printendnotes

%
% ==========   Bibliography
%
% \nocite{*}   % include everything in the uwthesis.bib file
\bibliographystyle{plain}
\bibliography{uwthesis}
%
% ==========   Appendices
%

\end{document}
